---
title: "PS270 Final Project"
author: "Cole Bakken"
description: "International Housing Prices and Working From Home"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction

This project examines the relationship between working from home and housing prices across several countries. Specifically, it asks whether higher levels of remote work are associated with higher real housing prices.

This question is important because the rise of remote work may change how people choose where to live. If individuals are less tied to physical workplaces, they may demand larger homes or relocate to different areas, potentially driving up housing prices. I hypothesize that countries with higher levels of working from home will have higher housing prices, as increased flexibility in work location raises demand for housing. I also hypothesize that if current oil prices stay elevated, this may also drive demand for working from home long-term, as well. 

The working from home data is sourced from: Aksoy, Cevat Giray, Jose Maria Barrero, Nicholas Bloom, Steven J. Davis, Mathias Dolls and Pablo Zarate, 2022. “Working from Home Around the World,” Brookings Papers on Economic Activity. Real Housing Prices are sourced from the Federal Reserve Bank of Dallas. 

```{r, echo = FALSE}
library(tidyverse)
library(readxl)
library(zoo)
library(knitr)
```

```{r, echo= FALSE}
rhpi <- read_excel("data/hp2503.xlsx", sheet = "RHPI")
```

```{r, echo = FALSE}
rhpi <- rhpi[-1, ]
rhpi$`...28` <- NULL
names(rhpi)[1] <- "Date"
```

```{r , echo = FALSE}
rhpi$Date <- as.yearqtr(rhpi$Date, format = "%Y:Q%q")

rhpi <- rhpi |>
  mutate(across(-Date, as.numeric))
```

```{r , echo = FALSE}
rhpi_longer <- rhpi |>
  select(Date, US, UK, Australia, France, Germany, Japan, Italy, Sweden, Spain, Canada) |> 
  pivot_longer(
    cols = -Date,
    names_to = "Country",
    values_to = "RHPI"
  )
```

## Data Visualization

```{r, echo = F}
ggplot(rhpi_longer, aes(x = Date, y = RHPI, color = Country)) +
  geom_line() + 
  labs (
    title = "Real House Price Index Over Time",
    x = "Year",
    y = "Real House Price Index (RHPI)",
    color = "Country"
  ) 
```
The real house price index is "seasonally adjusted and re-based" so the year 2005 equals 100 (from the Fed's Methodological Note). That means all countries being scaled to this reference point allows us to compare how house prices have changed over time across countries. As a result, the lines in the graph move relative to increases/decreases from the base value rather than representing raw price values. 

The graph shows that Canada has seemingly had the greatest increase in real house prices since 2000 while some countries experience much lower price levels. This variation suggests that there is meaningful cross-country difference in the dependent variable, which is important for testing the hypothesis of WFH's impact.

To evaluate the hypothesis, I examine whether there is a positive relationship between working from home and real housing prices. Evidence supporting the hypothesis would consist of a positive association between the average number of days worked from home and the real housing price index. In the regression analysis, this would correspond to a positive coefficient on the working-from-home variable. If the coefficient is close to zero or negative, this would provide evidence against the hypothesis.

```{r, echo = FALSE}
wfh <- read_excel("data/G-SWA.xlsx")
```

```{r, echo = FALSE}
wfh <- wfh |>
  mutate(year = case_when(
    wave == 1 ~ 2021,
    wave == 2 ~ 2022
  ))
```

```{r, echo = FALSE}
wfh_countries <- wfh |>
  filter(original_country %in% c("Australia", "France", "Germany", "USA", "UK", "Japan", "Italy", "Sweden", "Spain", "Canada"))

wfh_countries <- wfh_countries |>
  mutate(original_country = case_when(
    original_country == "USA" ~ "US",
    TRUE ~ original_country  ## This line will keep all the rest the same
  ))
```

```{r, echo = FALSE}
wfh_countries_mean <- wfh_countries |>
  group_by(original_country, year) |>
  summarize(
    wfh_mean = mean(n_work_home, na.rm = T),
   .groups = "drop" ## Tibble is still grouped by original_country otherwise 
   )
```

```{r, echo = FALSE}
rhpi_yearly <- rhpi_longer |>
  mutate(year = as.numeric(format(Date, "%Y"))) |> # Had to look up how to do this, don't believe it's Zoo, but base R
  group_by(Country, year) |>
  summarize(
    RHPI = mean(RHPI, na.rm = T,
    .groups = "drop"
    )) |>
  filter(year %in% c(2021, 2022))
```

```{r, echo = FALSE} 
merged_data <- rhpi_yearly |>
  inner_join(wfh_countries_mean, by = c("Country" = "original_country", "year" = "year"))
```

## Result Visualization 
```{r, echo = FALSE}
ggplot(merged_data, aes(x = wfh_mean, y = RHPI, color = Country)) +
  geom_point(size = 1.9) +
  geom_path(aes(group = Country), 
arrow = arrow(type = "open", length = unit(0.09, "inches"))) +
  geom_smooth(method = "lm", se = T, color = "black", alpha = 0.1) +
  guides(color = guide_legend(override.aes = list(linetype = 0))) + ## The arrows were being added to the country guide on the right, this is to prevent that. 
  labs(
    title = "WFH vs Real Housing Prices (2021-2022)",
    x = "Average Days Worked From Home in a Week",
    y = "Real Housing Price Index (2005 Base Year)",
    caption = "Data: US Dallas Fed & Brookings WFH Data"
  )
```
The graph shows the relationship between the average days worked from home and the real housing price index across countries in 2021 and 2022. Each country is represented by one or two points, depending on available survey data. This allows for cross-country differences and yearly changes for some of the countries. The plot shows a generally positive relationship between working from home and real housing prices across countries in 2021–2022, as indicated by the upward-sloping regression line. Countries with higher average days worked from home tend to have higher values of the real house price index (RHPI), although there is substantial variation around this trend.

## Regression Analysis 
```{r, echo = FALSE}
regression <- lm(RHPI ~ wfh_mean + factor(year), data = merged_data)

library(broom)

kable(
  tidy(regression), ## tidy(), and broom to turn it into a table was suggested when promting AI after I had difficulties using model summary
  digits = 3,
  col.names = c("Term", "Estimate", "Std. Error", "t value", "p-value"))
```

The regression results reinforce the plot's findings but indicate that the relationship is not statistically strong. The coefficient on working from home is positive (44.408), meaning that an increase of one additional day worked from home per week is associated with an increase of about 44 points in the RHPI, holding year constant. However, this estimate is not statistically significant (p = 0.13), meaning we cannot rule out that the observed relationship is due to random chance in the sample. The coefficient on the year indicator suggests that housing prices were slightly lower in 2022 compared to 2021, but this effect is very small and highly insignificant (p = 0.80). We should not take a causal interpretation of the results.

## Conclusion

Overall, this data analysis project provides limited support for the hypothesis that higher levels of working from home are associated with higher housing prices. While both the graph and regression results suggest a positive relationship, the estimated effect is not statistically significant and the model explains only a small portion of the variation in housing prices. There are several limitations to the analysis, including the very small sample size, the short time period, and the omission of important (possibly confounding) variables such as interest rates, income levels, and housing supply conditions, etc. 

Thus, we cannot take a causal interpretation of the results. 

With more time and money, the analysis could be improved by incorporating additional years, expanding the set of control variables as mentioned above, or the usage of different methods, instead of relying on survey data alone for WFH data. 