---
title: "PS270 Final Project"
author: "Cole Bakken"
description: "International Housing Prices and Working From Home"
output:
  distill::distill_article:
    self_contained: false
---

## Load & Clean Housing Data

```{r}
library(tidyverse)
library(readxl)
library(zoo)
```

```{r}
rhpi <- read_excel("data/hp2503.xlsx", sheet = "RHPI")
```

```{r}
rhpi <- rhpi[-1, ]
rhpi$`...28` <- NULL
names(rhpi)[1] <- "Date"
head(rhpi)
```

Here I deleted the first row of the data set, as it is an empty row in the excel file, as well as with the blank 28th column. I also renamed the 1st column to "Date". 



## Changing the Date Format of Real Housing Data

```{r}
rhpi$Date <- as.yearqtr(rhpi$Date, format = "%Y:Q%q")

rhpi <- rhpi |>
  mutate(across(-Date, as.numeric))

head(rhpi)
```

```{r}
rhpi_longer <- rhpi |>
  select(Date, US, UK, Australia, France, Germany, Japan, Italy, Sweden, Spain, Canada) |> 
  pivot_longer(
    cols = -Date,
    names_to = "Country",
    values_to = "RHPI"
  )

head(rhpi_longer)
```

The date format of this data is usual, and comes in the format of 1975:Q1, 1975:Q2. I saw a package online called "ihpdr" that already did the cleaning of the data, which I think would defeat the purpose of this project, but I noticed that the dates were changed to 1975-01-01, 1975-04-01, etc. 

After prompting AI, it recommended me to use the package "zoo" to work with the unusual time format for myself. The following code is written with the help of AI as I am unfamiliar how to work with zoo or clean data very well. It was also recommended to reshape the data longer (rather than wider as we have done in class)


## Data Visualization

```{r}
ggplot(rhpi_longer, aes(x = Date, y = RHPI, color = Country)) +
  geom_line() + 
  labs (
    title = "Real House Price Index Over Time",
    x = "Year",
    y = "Real House Price Index (RHPI)",
    color = "Country"
  ) 
```
The real house price index is "seasonally adjusted and re-based" so the year 2005 equals 100 (from the Fed's Methodological Note). That means all countries being scaled to this reference point allows us to compare how house prices have changed over time across countries. As a result, the lines in the graph move relative to increases/decreases from the base value rather than representing raw price values. 

The graph shows that Canada has seemingly had the greatest increase in real house prices since 2005, and in Milestone 4, I would hope to use working from home data to see any associations between real housing prices and the rate of working from home. My hypothesis is that Canada would have the highest rate compared to the US or the UK, as they have lower real home prices. 

## Load Work From Home Data

Source: Aksoy, Cevat Giray, Jose Maria Barrero, Nicholas Bloom, Steven J. Davis, Mathias Dolls and Pablo Zarate, 2022. “Working from Home Around the World,” Brookings Papers on Economic Activity.

```{r}
wfh <- read_excel("data/G-SWA.xlsx")
```

## Adding Year to Working From Home Data

```{r}
wfh <- wfh |>
  mutate(year = case_when(
    wave == 1 ~ 2021,
    wave == 2 ~ 2022
  ))
```

## Filtering Countries Down & Matching Names

```{r}
wfh_countries <- wfh |>
  filter(original_country %in% c("Australia", "France", "Germany", "USA", "UK", "Japan", "Italy", "Sweden", "Spain", "Canada"))

wfh_countries <- wfh_countries |>
  mutate(original_country = case_when(
    original_country == "USA" ~ "US",
    TRUE ~ original_country                         ## This line will keep all the rest the same
  ))
```

## Average Days Working From Home by Year 

```{r}
wfh_countries_mean <- wfh_countries |>
  group_by(original_country, year) |>
  summarize(
    wfh_mean = mean(n_work_home, na.rm = T),
   .groups = "drop" ## Tibble is still grouped by original_country otherwise 
   )
```

## Creating a yearly Real Housing Price, to compare with WFH data, then filtering by 2021/22
```{r}
rhpi_yearly <- rhpi_longer |>
  mutate(year = as.numeric(format(Date, "%Y"))) |> # Had to look up how to do this, don't believe it's Zoo, but base R
  group_by(Country, year) |>
  summarize(
    RHPI = mean(RHPI, na.rm = T,
    .groups = "drop"
    )) |>
  filter(year %in% c(2021, 2022))
```

## MERGE THE DATA! 
```{r}
merged_data <- rhpi_yearly |>
  inner_join(wfh_countries_mean, by = c("Country" = "original_country", "year" = "year"))

merged_data
```

## Quick Visualization 
```{r}
ggplot(merged_data, aes(x = wfh_mean, y = RHPI, color = Country)) +
  geom_point(size = 1.9) +
  geom_path(aes(group = Country), 
arrow = arrow(type = "open", length = unit(0.09, "inches"))) +
  geom_smooth(method = "lm", se = T, color = "black", alpha = 0.1) +
  guides(color = guide_legend(override.aes = list(linetype = 0))) + ## The arrows were being added to the country guide on the right, this is to prevent that. 
  labs(
    title = "WFH vs Real Housing Prices (2021-2022)",
    x = "Average Days Worked From Home in a Week",
    y = "Real Housing Price Index (2005 Base Year)",
    caption = "Data: US Dallas Fed & Brookings WFH Data"
  )
```
The graph shows the relationship between the average days worked from home and the real housing price index across countries in 2021 and 2022. Each country is represented by one or two points, depending on available survey data. This allows for cross-country differences and yearly changes for some of the countries.

There appears to be a positive relationship between working from home and housing prices, as indicated by the line of best fit, countries with higher levels of remote work like Canada and Australia tend to have higher real housing prices than the lower countries like Japan. However, there are some cases where from 2021 to 2022, housing prices increased despite working from home decreasing, showing that there are other factors influencing housing prices as well. 

The analysis is limited and a relatively short time frame. The results should be interpreted as descriptive rather than causal. 
