---
title: "PS270 Final Project"
author: "Cole Bakken"
description: "International Housing Prices and Working From Home"
output:
  distill::distill_article:
    self_contained: false
---

## Load & Clean Housing Data

```{r}
library(tidyverse)
library(readxl)
library(zoo)
```

```{r}
rhpi <- read_excel("data/hp2503.xlsx", sheet = "RHPI")
```

Here I deleted the first row of the data set, as it is an empty row in the excel file, as well as with the blank 28th column. I also renamed the 1st column to "Date". 

```{r}
rhpi <- rhpi[-1, ]
rhpi$`...28` <- NULL
names(rhpi)[1] <- "Date"
head(rhpi)
```

The date format of this data is usual, and comes in the format of 1975:Q1, 1975:Q2. I saw a package online called "ihpdr" that already did the cleaning of the data, which I think would defeat the purpose of this project, but I noticed that the dates were changed to 1975-01-01, 1975-04-01, etc. 

After prompting AI, it recommended me to use the package "zoo" to work with the unusual time format for myself. The following code is written with the help of AI as I am unfamiliar how to work with zoo or clean data very well. It was also recommended to reshape the data longer (rather than wider as we have done in class)

```{r}
rhpi$Date <- as.yearqtr(rhpi$Date, format = "%Y:Q%q")

rhpi <- rhpi |>
  mutate(across(-Date, as.numeric))

head(rhpi)
```

```{r}
rhpi_longer <- rhpi |>
  select(Date, US, UK, Canada, Australia) |> # this can be more countries or all later, but for this milestone, we will just use 4.
  pivot_longer(
    cols = -Date,
    names_to = "Country",
    values_to = "RHPI"
  )

head(rhpi_longer)
```

## Data Visualization

```{r}
ggplot(rhpi_longer, aes(x = Date, y = RHPI, color = Country)) +
  geom_line() + 
  labs (
    title = "Real House Price Index Over Time",
    x = "Year",
    y = "Real House Price Index (RHPI)",
    color = "Country"
  ) 
```
The real house price index is "seasonally adjusted and re-based" so the year 2005 equals 100 (from the Fed's Methodological Note). That means all countries being scaled to this reference point allows us to compare how house prices have changed over time across countries. As a result, the lines in the graph move relative to increases/decreases from the base value rather than representing raw price values. 

The graph shows that Canada has seemingly had the greatest increase in real house prices since 2005, and in Milestone 4, I would hope to use working from home data to see any associations between real housing prices and the rate of working from home. My hypothesis is that Canada would have the highest rate compared to the US or the UK, as they have lower real home prices. 

## Load Work From Home Data

```{r}
wfh <- read_excel("data/G-SWA.xlsx")
head(wfh)
```

This data set has 36k obs. so it will take me a bit to figure out how to use it and filter it the best, but that is for future me! This is data only from 2021-2022, and I don't believe I can access any more years. 

Source: Aksoy, Cevat Giray, Jose Maria Barrero, Nicholas Bloom, Steven J. Davis, Mathias Dolls and Pablo Zarate, 2022. “Working from Home Around the World,” Brookings Papers on Economic Activity.